---
title: "HW01_rishabk2"
author: "Rishab Kulkarni"
date: "1/18/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1a.
```{r}
set.seed(678461346)

x <- rnorm(100, mean = 3, sd = 2)
x[1:10]
```

Set seed with my UIN as the seed; Then, I used the rnorm() function with mean=3
and sd=2 as parameters to produce 100 values. Stored these values in a vector x,
and printed out the first 10 values.

# 1b.

The formula for sample mean is:

$\frac{1}{n}\ \sum_{i=1}^{n} x_i$

The formula for sample variance is:

$\frac{1}{n-1}\ \sum_{i=1}^{n} (x_i-\bar{x})^2$

# 1c.
```{r}
mysummarystat <- function(x){
 x_b <- sum(x) / length(x)
 x_v <- sum((x - x_b)^2) / (length(x)-1)
 
 return(c(x_b, x_v))
}
mysummarystat(x)
```

Found the sample mean by dividing the sum of the values by 100, or number of
values. Found the sample variance by dividing the rss by number of values-1.
The function returns a vector containing these two values.

```{r}
mean(x)
```

```{r}
var(x)
```

The R functions mean() and var() output nearly the same values computed by
the function.

# 2a.
```{r}
data("iris")

# changing class labels
levels(iris$Species) <- c('Species_1', 'Species_2', 'Species_3')

# showing new class labels
iris$Species
```

# 2b.
```{r}
table(iris$Species)/nrow(iris)
```

The frequencies for each species are the same: 1/3

# 2c.
```{r}
cont <- iris[, -5]

cor(cont)
```

cont includes only continuous variables, excluding Species. Used the cor() 
function with cont as the parameter to produce the correlation matrix.

Petal.Width and Petal.Length have the highest correlation of ~0.963.

# 2d.
```{r}
e <- eigen(cor(cont))
e
```

Used the eigen() function to compute eigenvalues and eigenvectors of the
correlation matrix found in part 2c. 

The largest eigenvalue is 2.91849782, and the corresponding eigenvector is
(0.5210659, -0.2693474, 0.5804131, 0.5648565).

# 3a.
```{r}
boxplot(cont)
points(colMeans(cont),pch = 19,
       cex = 1.5, col = "red")
```

# 3b.
```{r}
boxplot(cont)
points(colMeans(cont),pch = 19,
       cex = 1.5, col = "red")

# box-plot adjustments
fig.align = 'center'
out.width = "75%"
```

Centered box plot with page width of 75%.

# 4a.
```{r}
install.packages("devtools",repos = "http://cran.us.r-project.org")

library(devtools)
install_version("ElemStatLearn", version = "2015.6.26.2", 
                  repos = "http://cran.us.r-project.org")

library(ElemStatLearn)
data("ozone")

# saving ozone data into a .csv file
write.csv(ozone, 'ozone.csv')

# read data from .csv file back into R
ozone_csv <- read.csv(file = 'ozone.csv',row.names = 1)

head(ozone_csv)
```

First column of ozone_csv was simply an indexing column, but the original ozone
data didn't have this. Thus, it was excluded when printing out the first 5 
observations via row.names = 1.
